---
title: "Another way to use wearable computing devices"
author: "Patrick Alex Freitas da Silva - patrickalex@gmail.com"
date: "january 13th, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The use of data from accelerometers on belt, forearm, arm and dumbell of 6 participants in order to evaluate the quality of the activities that had been done.

## Objective

Use data and machine learning techniques in order to predict which activities had been done

## How the model was built

## How cross validation was used

## Prognosis about the expected out of sample error

## Reasons for the choices made

## Prediction Assignment
#### About the data set
The data for this project came from 

#### Loading the data training and testing set
```{r loadingDataSets}
trainingSet <- read.csv('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv')
testSet     <- read.csv('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv')
```

####


#### References

Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6.



[2]:http://groupware.les.inf.puc-rio.br/har#ixzz4VdrEh3GS
[1]:http://groupware.les.inf.puc-rio.br/har